RBE/CS549 Project: Boat Tracking/Detection

### 2017-12-12 paper/main.tex
Worked on optical flow section

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{pdfpages}
\usepackage{minted}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\graphicspath{ {images/} }
\title{RBE/CS549 Project: Boat Tracking/Detection}
\author{Jordan Burklund \and James Kuszmaul}
\begin{document}
\maketitle

@@

@@

As an overview, the way that this optical flow algorithm works is to

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{pdfpages}
\usepackage{minted}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\graphicspath{ {images/} }
\title{RBE/CS549 Project: Boat Tracking/Detection}
\author{Jordan Burklund \and James Kuszmaul}
\begin{document}
\maketitle

@@

\subsection{Underlying Algorithm}

@@

As an overview, the way that this optical flow algorithm works is to
approximate the neighborhood of every pixel as some quadratic function
$f(\mathbf{x}) = \mathbf{x}^T \mathbf{A} \mathbf{x} + \mathbf{b}^T \mathbf{x} + c$
where $\mathbf{x}$ is the location around any given pixel. By estimating these
polynomials about any given pixel between two frames, we can try to estimate how
much the image has moved in the neighborhood of that pixel. Actual
implementations of the algorithm contain some additional filtering/smoothing and
function on multiple levels of an image pyramid to estimate optical flow at
different scales. You can also adjust how much of the neighborhood around a
pixel is used to estimate the polynomial, and various other optimizations.

In practice, we just used the default options to the matlab method (which uses
a 3-level pyramid scaling image size by a factor of 2, with a 5-pixel
neighborhood size and a Gaussian filter on the output of size 15x15), as these
options tended to produce good results.

\subsection{Expected Flow}

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{example_flow.png}
\caption{Optical Flow output (shown only for every few pixels) between two
frames}
\label{fig:example_flow}
\end{figure}

In Figure \ref{fig:example_flow} we show the output of
\texttt{opticalFlowFarneback} on a given pair of frames. The most immediate need
in this image is to try and account for the field of flow produced along the
water. This corresponds to the real movement of the plane of the water relative
to the boat. We can attempt to model this with the following assumptions:
\begin{itemize}
\item That the camera is pointed straight and level
\item There is no substantial roll or yaw angular rates of turn occurring
\item The camera is an ideal pinhole camera with no substantial lens distortions
      or the such
\item There is no sideways relative velocity between the boat and the water
\end{itemize}

We can define the following variables:

\begin{tabular}{c|c}
Name & Description \\ \hline
$h$ & Height of the camera above the water \\
$v_h$ & The current upwards velocity of the camera \\
$v_b$ & Forwards velocity of the boat relative to the water \\
$x_b, y_b$ & Position coordinates in the boat frame \\
$x_f, y_f$ & X and Y position in the 2-D image frame \\
$\dot{x}_f, \dot{y}_f$ & The expected optical flow (derivatives of $x_f, y_f$)
\\
$K$ & Some constant describing the field of view of the camera \\
\end{tabular}

The boat frame has an origin immediately below the camera on the surface of
the water, is attached to the boat, has an x-axis pointing straight forwards,
a y-axis pointing straight port (left in the camera view) and a z-axis pointing
straight upwards, with the camera at position $(0, 0, h)$ in this frame.

The image frame is scaled so that the center of the image is the origin and
the image is exactly 1 unit tall, with the x-axis being the side-to-side axis
and the y-axis being up-and-down.

Strictly speaking $K = 2\sin\frac\theta2$ if the total vertical field-of-view
of the camera is $\theta$, but we end up estimating it anyways.

These give us:

\begin{align*}
x_f =& \frac{-y_b}{Kx_b} \\
y_f =& \frac{-h}{Kx_b} \\
\dot{x}_f =& \frac{-y_bv_b}{Kx^2_b} \\
\dot{y}_f =& -\frac{1}{Kx_b}(\frac{hv_b}{x_b} + v_h) \\
\end{align*}

And solving for $x_b, y_b$ in terms of $x_f, y_f$ and then substituting in to
give us the expected optical flow at any given image coordinate $x_f, y_f$ we
get:

\begin{align}
\dot{x}_f =& \frac{-v_bKx_fy_f}{h} \\
\dot{y}_f =& \frac{v_hy_f - Kv_by_f^2}{h}
\end{align}

If we ignore the actual values of the constants and just turn them into
something to be estimated, while also adding in a parameter where we
assume that $y_f$ may not be perfectly centered, we can turn this into a
linear parameter estimation problem as:

\begin{align}
\dot{x}_f =& ax_fy_f + bx_f \\
\dot{y}_f =& cy_f^2 + dy_f + e
\end{align}

To actually perform the least-squares fitting we use \texttt{fitlm} in
matlab. Normally, this just uses a normal least-squares estimation
(similar to doing \texttt{A \ b}), but adds some additional
options. We tried to make use of these options, including a outlier
rejection algorithm, but the effects were minimal and substantially
increased the processing time required to execute the code.

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{diff_magnitude}
\caption{Difference between the measured and modelled optical flow.
         Dark blue means perfect match, bright yellow means a poor match.}
\label{fig:diff_magnitude}
\end{figure}

When we compare the expected magnitude of the optical flow to the
actual magnitude of the optical flow in Figure \ref{fig:diff_magnitude}
(this is the same image as Figure \ref{fig:example_flow}),
in which you can see a few features:
\begin{itemize}
\item The boat in the left half of the image shows up as a cyan blob, distinct
      from the water around it.
\item The curved metal in the bottom left screws with the optical flow
      algorithm, causing substantial variation from the model.
\item In the bottom right, due to either lens distortion or hard to measure
      waves, the optical flow varies substantially from the model.
\end{itemize}

As a result of the poor results in the very bottom of the image and in the
area around the metal bar, we completely ignore and filter out these regions,
as they are relatively unlikely to contain obstacles (or if they did, our
boat would be running them over anyways).

We also note that in general objects appearing more distantly in the center of
the image will be hard to identify, as the absolute value of the difference
in magnitude from the modelled magnitude will be low (because we are looking
at optical flow in the image, not real life, so far away objects have
low velocities in the image), and so simple thresholds will not work well.
We experimented with reducing the required threshold in the center of the image,
which did improve tracking.

It is also the case that any stationary, low-lying obstacles (e.g. buoys)
will be indistinguishable from the water and so will not be detected
by this half of the filtering.

\section{Blob Detection and Filtering}

\section{Object Matching}

\section{Results}

\section{Future Results}
